% Released under the MIT license.

\documentclass{ronr-bylaws}

\newcommand{\refsection}[2]{Article \ref{#1} \S \ref{#2}}
\newcommand{\refarticle}[1]{Article \ref{#1}}

\author{Nicolas Nytko}

\partyname{University of Illinois Urbana-Champaign Chapter of Society for Industrial and Applied Mathematics}

\date{\today}

\begin{document}
%
%
\maketitlepage
%
%
\tableofcontents
\newpage
%
\maketitle
This Rules of Procedure (hereinafter called ``Rules'') apply to the Society for Industrial and Applied Mathematics (``SIAM'') Student Chapter, named the ``University of Illinois Urbana-Champaign Chapter of Society for Industrial and Applied Mathematics''.  Hereinafter, ``SIAM'' or ``Society for Industrial and Applied Mathematics'' will refer to the parent organization, whilst ``the Chapter'' will refer to the student chapter/organization.

The Chapter to which these Rules apply is formed by the Society for Industrial and Applied
Mathematics and shall operate within the Bylaws of the parent organization. The SIAM bylaws
specify how Chapters are formed; see the SIAM bylaws for details. The Chapter shall not affiliate
with any other organization without first obtaining the written approval of SIAM. Provisions for
SIAM  Student (University) Chapters are contained in the SIAM Bylaws and are included in these
Rules. No provisions of these rules shall be construed so as to contradict the Bylaws of SIAM.
\article{Purpose}
\section{SIAM Objectives}
The objectives of SIAM, as established in the Certificate of Incorporation, are:
\begin{itemize}
\item To further the application of mathematics to industry and science.
\item To promote basic research in mathematics leading to new methods and techniques useful to industry and science.
\item To provide media for the exchange of information and ideas between mathematicians and other technical and scientific personnel.
\end{itemize}
\section{Chapter Purposes}
Purposes of the Chapter shall be consistent with the objectives of SIAM. The specific purposes of the Chapter are:
\begin{itemize}
\item To create a community for students interested in applied and computational mathematics.
\item To promote applied mathematics and computational science to both graduate and undergraduate students throughout the computer science department and across campus.
\end{itemize}
\article{Activities}
The Chapter shall schedule several types of activities. Possible activities include:
\begin{itemize}
\item Seminars with themes of interest to students from different fields.
\item Short talks on a variety of topics such as overviews of speakers' research, surveys of the research in a particular field, interesting problems, career options and opportunities talks, and talks from recruiters.
\item Social activities aimed at strengthening the community served by the Chapter.
\end{itemize}
\article{Institution Served}
Members  of the Chapter will be recruited from the University of Illinois Urbana-Champaign.
\article{Membership}
\section{Elegibility}
Any person engaged or interested in mathematics or computing and their applications
shall be eligible for membership in this Chapter. Chapter membership may be interdisciplinary,
with members from multiple departments.

All members of the Chapter who are not already affiliated with the parent organization shall be
encouraged to become regular or student members of SIAM.

\section{Member classes}
The Chapter shall have two classes of members: voting and non-voting. All Chapter
members who are members of SIAM are eligible to vote in Chapter elections and referenda, provided
they have been members of the Chapter for at least thirty days. Members of the Chapter who are
not members of SIAM are ineligible to vote in Chapter elections and referenda.

\section{Membership fees}
Membership in the Chapter shall be free of charge. However, the Chapter may levy
fees to cover the expenses of special activities, such as social activities and trips to labs, but the
collection of fees is to be avoided, whenever possible.

\section{SIAM membership}
Chapter members shall have the privileges of SIAM membership only if they are regular
or student members of SIAM.

\section{SIAM student membership}
All members of the Chapter who are students enrolled in the sponsoring institution are
eligible for free student memberships in SIAM. The Chapter is responsible for providing a list of its
student members to SIAM so that complimentary student membership in SIAM can be processed.

\section{Termination of membership}
Termination of student membership shall take place upon graduation or withdrawal
from the University of Illinois Urbana-Champaign.
%
\article{Sponsorship}
\section{Sponsor}
The Chapter is a registered student organization (RSO) of the University of Illinois Urbana-Champaign.  The Chapter and its operations are not, however, directly sponsored by the university.

\section{Faculty Advisor}
The executive committee of the chapter shall appoint a Faculty Advisor for the Chapter.  The Advisor must be a member in good standing with SIAM. In the event said Advisor relinquishes his/her position, a new Advisor shall be appointed.  The responsibilities, rights and duties of the Faculty Advisor shall be those normally assigned to the Faculty Advisor of student organizations of the University, but in addition, the Faculty Advisor is expected to take leading role in the development of the Chapter activities consistent with the objectives of SIAM.
%
\article{Officers}\label{article:officers}
\section{Officer Roles}\label{section:officer_roles}
The Chapter shall have a President, a Vice-President, a Secretary and a Treasurer. Officers shall be chosen from Student Members of the Chapter and must  be Student Members in good standing with SIAM.

\section{President}\label{section:president}
The President shall preside at the meetings of the Chapter (and the Chapter Executive Committee, see \refarticle{article:executive_committee}).  In the absence of the President, the Vice-President shall assume the duties of the President.  In the absence of the latter, the Secretary shall assume said duties.

\section{Secretary}\label{section:secretary}
The Secretary shall keep a record of the affairs of the Chapter, handle correspondence, and submit an annual report of Chapter activities to the Secretary of SIAM, which report shall be suitable for publication in SIAM News or its equivalent.

\section{Treasurer}\label{section:treasurer}
The Treasurer shall receive and take custody of Chapter funds, and shall submit an annual Treasurer's Report and other financial reports, as requested, to the Treasurer of SIAM.  The annual Treasurer's Report shall be prepared as of the end of the academic year and shall be transmitted to the Treasurer of SIAM by no later than 30 days following the end of the academic year.

\section{Officer Elections}\label{section:officer_elections}
Officer elections shall be held at a regularly scheduled meeting before the end of the
Spring Semester. The date of the election shall be announced to all members no less than thirty
days in advance. Early voting and voting by proxy may be permissible, given that appropriate
steps to ensure accountability are taken.

Nominations for officer candidates shall be finalized no less than one week prior to the vote. The
Chapter Secretary shall be responsible to verify all nominations are in compliance with the terms
in \refsection{article:officers}{section:officer_roles}.

The election shall be held by secret ballot prepared by the Chapter Secretary in advance. The
Secretary shall also be responsible for tallying the votes and determining winners. The vote count
shall be open and may be observed by any Chapter member.

The elected term shall begin on the following August 1 and last through July 31 of the next year.

\section{Officer Terms}
Each officer term shall last one year. An officer may be reelected to the same position,
but no position shall be held by the same individual more than two consecutive years.

\section{Early Vacation}\label{section:early_vacation}
In the event that an officer vacates their position early, an election shall be held at the next
scheduled meeting to select an immediate replacement. Any of the current officers may elect to
run for the newly opened position. In the event that an officer is to fill the vacant seat, another
election must be held to fill that officer’s newly vacated seat. Time served in a vacated position
shall not count toward the maximum term limit.

\article{Executive Committee}\label{article:executive_committee}
\section{Committee members}
The Chapter shall be managed by an Executive Committee consisting of the incumbent
officers and a single member appointed by the Faculty Advisor. Any active student member of the
Chapter may be appointed to this position, provided they are not currently serving as an officer.

The Advisor-appointed Committee member  may be re-appointed and may serve a maximum  of
three consecutive terms.

The Executive Committee is charged with making policy decisions for the Chapter. The Chapter
President shall chair the Executive Committee, and the Chapter Secretary shall keep records of
meetings and decisions made by the Executive Committee.

\section{Vacant seats}
Should a seat in the Executive Committee be vacated prematurely by one of the officers,
it shall be filled following the rules in \refsection{article:officers}{section:early_vacation}. If the Advisor-appointed member steps down, the Advisor shall appoint a new member to the committee to serve the remainder of the term. Time served in a vacated position shall not count toward the maximum term limit.

\article{Other Committees}
\section{Creation of Committees}
To provide the Chapter with flexibility to adapt, the Executive Committee may easily
create and dissolve committees, with the exception of itself. The official creation of a committee
may be done by the President, Vice-President, or by a majority vote by the Executive Committee.
The official dispersement of a committee requires majority agreement in the Executive Committee.

A committee may be organized in whatever way is deemed best to achieve its goals. The Executive
Committee may  appoint individuals to positions in a committee.

Examples of possible committees include:
\begin{itemize}
  \item An activities committee responsible for suggesting Chapter activities and investigating the
    feasibility of those activities.
  \item A committee designed to work on an upcoming event, such as Math Awareness Month activities.
\end{itemize}

The purpose of committees is to distribute responsibility to Chapter members and to give members
an official capacity to influence the Chapter without being in the Executive Committee.

\section{Nominating Committee}
In most cases, informal nominations, including self-nominations, for officer position
candidates shall be sufficient. However, a nominating committee may be appointed by the Executive
Committee at its discretion. The design of this committee shall also be at the discretion of the
Executive Committee. To ensure fairness, members may be nominated outside of the committee by
gathering support for their nomination by petition. A successful petition must contain signatures
from at least ten percent of the Chapter's voting members. Should the Chapter have more than
300 voting members, a successful petition need only have thirty signatures.

Any officer nominees must be eligible as stated in \refarticle{article:officers}. Regardless of the method of nomination, elections shall be conducted as detailed in \refsection{article:officers}{section:officer_elections}.

\article{Meetings}
\section{Minimum meeting count}
There shall be a minimum of five meetings per year. Meetings times shall be announced
to all Chapter Members at least two weeks in advance. Topics related to Chapter business and
policy will be selected by the Officers in advance of the meeting, and time should be provided for
unplanned topics to be discussed.

Meetings shall be presided over as stated in \refsection{article:officers}{section:president}.

\section{Business meetings}
The Chapter shall conduct a business meeting once per year during the month of
August. Other business meetings may be called by the President or the Treasurer with two weeks
advance notice.
%
\article{Chapter Funds}
\section{Membership fees}
The Chapter shall not charge membership fees. General funds shall be acquired from
sources such as the SIAM parent organization.

In some cases, the Chapter may levy fees to cover the expenses of special activities.

\section{Unused funds}
The Chapter shall deposit all unused funds to which it has legal title in excess of \$200
in an insured savings account, unless current operating commitments are in excess of that amount
or unless the Chapter Treasurer obtains a written authorization from the SIAM Treasurer.

\section{Books of account}
The Treasurer shall maintain books of account that show income and expense items
for all activities and balances for all accounts of the Chapter.

\section{Grants or Loans}
The Chapter may request a grant or loan from the Treasurer of SIAM under the provision of Article XII, Chapter 8 of the Bylaws of SIAM. Such requests shall be made by submission of
the ``Request for Funding'' form to SIAM and include a current financial statement for the Chapter
and a proposed budget for the requested funds.

\section{External funding}
Other than seeking funds from the parent institution of the Chapter, no officers
or member of the Chapter may apply for a grant to support the Chapter activities or enter into
any contract to support such activities or provide services, without approval of the President and
Treasurer of SIAM or the Executive Director acting on behalf of the Treasurer.

\article{Amendments}
\section{Alteration of rules}
These Rules may be altered or amended with the approval of the SIAM Board of Trustees. Submission to the Board of proposal alterations or amendments shall only be made after approval by majority vote of members of the Chapter present (or represented by proxy) at a scheduled meeting.

\section{Advance notice of vote}
A vote to amend the Rules shall be announced no less than thirty days in advance. Additionally,
the complete text of the amendment shall be made available no less than thirty days prior to the
vote.

\section{Meeting time of vote}
If feasible, the vote to amend the Rules shall be scheduled during a regular meeting. If a special
meeting is called for the vote, then the time must be agreed upon by a majority of the Executive
Committee.

\article{Termination of the Chapter}
\section{Self-termination}
A Chapter may terminate itself by the unanimous vote of the members of the Chapter
present (or represented by proxy) at a scheduled meeting, provided notice of the proposed termi-
nation and the meeting at which it is to be considered has been given to all Chapter members at
least thirty days in advance.

\section{Inactivity}
A Chapter may be terminated by the board if there has been no Chapter activity for
one year.

\section{Relegation of funds}
In the event that the Chapter terminates, the funds to which it has legal title shall
revert to the account of SIAM.
\end{document}
